---
title: "LandCoverChange: Pipeline"
output: html_document
date: "2023-05-08"
editor_options: 
  chunk_output_type: console
---

```{r init, include = FALSE}
knitr::opts_chunk$set(echo = T) 
```


## Setup

```{r setup, echo = FALSE}
# library(LandCoverChange)
source("~/Documents/LandCoverChange/R/LandCoverChange.R")
# devtools::install_github("EarthSystemDiagnostics/corit")
library(corit)
library(tidyverse)
library(sf); sf_use_s2(FALSE)

data_dir   <- "~/Documents/LandCoverChangeProject_data/"

# output_dir <- "~/Documents/LandCoverChangeOutput_vignette"
# create_subfolders(output_dir, default = TRUE)
```


## Study region and vegetation data

Define regions ('regMap' for land cover reconstruction and 'calibMap' for the calibration of vegetation cover to land cover), create Pollen Source Areas (PSAs) from their site centres and pollen source radius, and extract vegetation cover time series in all PSAs.

```{r psaMap, include=TRUE, message=FALSE, warning=FALSE}

# define map and projection
centre = c(175, 65) # center of study region, (lon, lat) in °
extent = 2000       # 1/2 diameter of the rectangle to cut around centre in equal area projection
proj   = sprintf("+proj=laea +lon_0=%f +lat_0=%f +ellps=sphere", centre[1], centre[2])
bbox   = st_sfc(st_point(c(0,0)), crs = proj) %>% st_buffer(extent*1000) %>% st_bbox() %>%
            st_as_sfc()

world  = rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%
  st_shift_longitude() %>% st_transform(proj) %>% st_make_valid()

# region subset
regMap = suppressWarnings(
  world %>% st_intersection(bbox) %>% st_buffer(50) %>% st_union()
)

# calibration subset
calibMap = suppressWarnings(
  world %>% st_intersection(bbox %>% st_buffer(extent * 1000)) %>%
  dplyr::select("continent") %>% rename(groups = continent) %>%
  group_by(groups) %>% summarize(geometry = st_union(st_buffer(geometry, 20)))
)

### for cluster process
library(parallel)
cl <- makeCluster(detectCores()/2)

# load data, filter pollen source areas in region and/or calibration subset
psaVeg <- make_psaVeg(veg_folder = file.path(data_dir, "data", "vegetation_cover"),
                      regMap = regMap, calibMap = calibMap, interpolate = TRUE, fc = 1/500, dt = 500, k = 5, cluster = NULL)

```

```{r psaVegPlot, fig.height = 10, fig.width = 10}
plot(psaVeg)
```

Plot vegetation time series for specific sites
(options: plot only n_taxa taxa with largest mean abundance)

```{r plotVegSite, fig.height = 8, fig.width = 12}
plotVegSite(psaVeg, ID = unique(psaVeg@vegetation$Dataset_ID)[100], n_taxa = 10, nrow = 2, ncol = 5)
```

# Calibration of the Vegetation Cover - Land Cover relationship to Modern Data

## Modern vegetation cover shares within each pollen source area

First, we define the time 'modern' for which summarize the vegetation data. Then we load the landcover metadata sheet that contains information on merging of landcover classes.

```{r veg_modern_shares}
modernVegetation <- psaVeg@vegetation %>% filter(Interp, Age <= 1000) %>%
                      dplyr::select(-c(Interp, Age)) %>%
                      group_by(Dataset_ID) %>%
                      summarise(across(everything(), \(x) mean(x, na.rm = TRUE))) %>% filter(!apply(., 1, function(x) all(is.nan(x[-1]))))

class_defs <- readxl::read_excel(glue::glue("{data_dir}settings/landcover_class_definitions.xlsx")) 
```

## Google Earth Engine options

In several functions below, there will be options to use Googel Earth Engine for faster calculation of e.g. landcover shares.
To use this option you have to first configure `rgee` on your device and get it running (this can be a bit tricky at times).

```{r rgee_init, message = FALSE, echo = FALSE}
# install.packages("rgee")
## or
# library(remotes)
# install_github("r-spatial/rgee")

library(rgee)
# ee_install()

ee_Initialize() ## this function needs to run without error to use GEE
```

## modern land cover shares within each pollen source area

```{r lc_modern_shares, fig.height = 6, fig.width = 9, message = FALSE}

lcovShares <- modernLC(psaVeg, psaIDs = unique(modernVegetation$Dataset_ID), class_defs, GEE = TRUE)
## can take a minute

## example plot of three ids
## TODO: match colors with color code from class_defs
lcovShares %>% filter(Dateset_ID%in%unique(Dateset_ID)[sample(1:nrow(.), 3)]) %>%
  pivot_longer(cols = -Dateset_ID, values_to = "Percentage") %>% mutate(Class_Code = as.numeric(name)) %>%
  left_join(class_defs %>% dplyr::select(Class_Code, Class_Label)) %>%
  ggplot(aes(x = '', y = Percentage, fill = Class_Label)) +
    geom_bar(stat="identity", width=1) +
    coord_polar("y", start=0) +
    facet_wrap(~Dateset_ID) +
    theme_light() +
    theme(legend.position = "top")
```

## Translate Vegetation to Landcover




## End

```{r}
stopCluster(cl)
```


